---
title: "ST_541_FinalProject"
author: "Collins Mills, Mary Strickland, Carter Lockwood, Steven Johnson"
date: "December 9th 2020"
output: pdf_document
---
# Data Insert and Cleaning
```{r setup, include=FALSE}
library(readr)
library(leaps)
library("stringr")
library(ggplot2)
library(car)
library(dplyr)

IMDb_movies <- read.csv(file = "IMDb movies.csv", header=TRUE)
ratings <- read.csv(file = "IMDb ratings.csv", header=TRUE)
cpi_rates <- read.csv(file = "cpi_table.csv", header=TRUE)
```
## Renaming and inserting columns
```{r}
names(ratings)<-str_replace_all(names(ratings), c("_" = "." , "," = "" ))
names(IMDb_movies)<-str_replace_all(names(IMDb_movies), c("_" = "." , "," = "" ))
#rename 'worlwide' to 'worldwide'
colnames(IMDb_movies)[19] <- "worldwide.gross.income"
#Inserting dummy variables to account for Movie Genres originally combined into one column.
IMDb_movies["IsItHorror"]<- 0
IMDb_movies["IsItRomance"] <- 0
IMDb_movies["IsItAction"] <- 0
IMDb_movies["IsItComedy"] <- 0
IMDb_movies["IsItDrama"] <- 0
IMDb_movies$IsItHorror[grep("Horror",IMDb_movies$genre)] <- 1
IMDb_movies$IsItRomance[grep("Romance",IMDb_movies$genre)] <- 1
IMDb_movies$IsItAction[grep("Action",IMDb_movies$genre)] <- 1
IMDb_movies$IsItComedy[grep("Comedy",IMDb_movies$genre)] <- 1
IMDb_movies$IsItDrama[grep("Drama",IMDb_movies$genre)] <- 1
IMDb_movies$IsItHorror<-as.factor(IMDb_movies$IsItHorror)
IMDb_movies$IsItRomance<-as.factor(IMDb_movies$IsItRomance)
IMDb_movies$IsItAction<-as.factor(IMDb_movies$IsItAction)
IMDb_movies$IsItComedy<-as.factor(IMDb_movies$IsItComedy)
IMDb_movies$IsItDrama<-as.factor(IMDb_movies$IsItDrama)
```

## Data Subsetting

```{r}
#To account for NA records and extracting only records where the movie was made in the USA. 
#This decision was made to focus analysis and minimize the dataset. 
budget <-IMDb_movies[!is.na(IMDb_movies$budget),]
usamovies<-budget[!is.na(budget$usa.gross.income),]
usamovies1<-usamovies[!is.na(usamovies$country),]
usamovies2<-usamovies1[grep("USA",usamovies1$country),]
#Inserting a currency column to account for only records where profit was measure in dollars.
# This eliminated the need to convert currency and created consistency in the dataset.
usamovies2$Currency <- substr(usamovies2$budget, start = 1, stop = 3)
for(value in 1:6605){
  if(substr(usamovies2$Currency[value], start = 1, stop = 1)=="$"){
    usamovies2$Currency[value] = "USD"
  }
}
USAMOVIES<-subset(usamovies2, usamovies2$Currency== "USD")
#Recreating budget and gross income columns to remove "$" and change type to numeric
USAMOVIES$budget <- substr(USAMOVIES$budget, start = 2, stop = 15)
USAMOVIES$usa.gross.income <- substr(USAMOVIES$usa.gross.income, start = 2, stop = 15)
USAMOVIES$worldwide.gross.income <- substr(USAMOVIES$worldwide.gross.income, start = 2, stop = 15)
USAMOVIES$budget <- as.numeric(USAMOVIES$budget)
USAMOVIES$usa.gross.income <- as.numeric(USAMOVIES$usa.gross.income)
USAMOVIES$worldwide.gross.income <- as.numeric(USAMOVIES$worldwide.gross.income)
#Creating of profit column under the assumption that all budget was used and world wide gross income is the best representation of money earned.
USAMOVIES$profit <-(USAMOVIES$worldwide.gross.income- USAMOVIES$budget)

#make year column into the index for cpi_rates dataframe
row.names(cpi_rates) <- cpi_rates$Year
cpi_rates[1] <- NULL
cpi_2020 <-  cpi_rates["2020", 1]


#Formula for inflated values: (original value * 2020 Consumer Price Index Value) / Consumer Price Index From Year of Release
for(val in 1:nrow(USAMOVIES)){
  USAMOVIES[val, "budg_infl"] <- (USAMOVIES[val, "budget"] * cpi_2020) / cpi_rates[toString(USAMOVIES[val, "year"]), "Avg"]
  USAMOVIES[val, "worldwide_gross_infl"] <- (USAMOVIES[val, "worldwide.gross.income"] * cpi_2020) / cpi_rates[toString(USAMOVIES[val, "year"]), "Avg"]
  USAMOVIES[val, "usa_gross_infl"] <- (USAMOVIES[val, "usa.gross.income"] * cpi_2020) / cpi_rates[toString(USAMOVIES[val, "year"]), "Avg"]
  USAMOVIES[val, "profit_infl"] <- (USAMOVIES[val, "profit"] * cpi_2020) / cpi_rates[toString(USAMOVIES[val, "year"]), "Avg"]
}
names(USAMOVIES)<-str_replace_all(names(USAMOVIES), c("_" = "." , "," = "" ))

summary(USAMOVIES)
```
# Successful Movies
## Getting familiar with the data: Examining Count, Budget, Profit, and Score Over The Years
```{r include= FALSE}
m <- read.csv("usamovies_num.csv")
#rename 'worlwide' to 'worldwide'
colnames(m)[22] <- "worldwide.gross.income"
cpi_rates <- read.csv("cpi_table.csv")
#make year column into the index for cpi_rates dataframe
row.names(cpi_rates) <- cpi_rates$Year
cpi_rates[1] <- NULL
cpi_2020 <-  cpi_rates["2020", 1]


#Formula for inflated values: (original value * 2020 Consumer Price Index Value) / Consumer Price Index From Year of Release
for(val in 1:nrow(m)){
  m[val, "budg_infl"] <- (m[val, "budget"] * cpi_2020) / cpi_rates[toString(m[val, "year"]), "Avg"]
  m[val, "worldwide_gross_infl"] <- (m[val, "worldwide.gross.income"] * cpi_2020) / cpi_rates[toString(m[val, "year"]), "Avg"]
  m[val, "usa_gross_infl"] <- (m[val, "usa.gross.income"] * cpi_2020) / cpi_rates[toString(m[val, "year"]), "Avg"]
  m[val, "profit_infl"] <- (m[val, "profit"] * cpi_2020) / cpi_rates[toString(m[val, "year"]), "Avg"]
}
```
```{r out.width="50%" out.height="50%"}
decades <- c(1910, 1920, 1930, 1940, 1950, 1960, 1970, 1980, 1990, 2000, 2010, 2020)
m$decade <- decades[findInterval(m$year, decades)]
mean_budg_by_dec <- aggregate(budg_infl ~ decade, data = m, mean)
mean_prof_by_dec <- aggregate(profit_infl ~ decade, data = m, mean)
mean_score_by_dec <- aggregate(avg.vote ~ decade, data = m, mean)
count_by_decade_plot<- barplot(table(m$decade), main="Movie Count By Decade", xlab="Decade", ylab="Number of Movies")

mean_score_plot <- barplot(mean_score_by_dec$avg.vote, main="Average Score By Decade", names.arg=mean_score_by_dec$decade, xlab = "Decade", ylab="Avg Score")
text(x=mean_score_plot, y=mean_score_by_dec$avg.vote, label=round(mean_score_by_dec$avg.vote, digits=2), pos=3, cex=0.8, col="red")

mean_budg_plot <- barplot(mean_budg_by_dec$budg_infl, main="Average Budget By Decade", names.arg=mean_budg_by_dec$decade, xlab = "Decade", ylab="Avg Budget")
text(x=mean_budg_plot, y=mean_budg_by_dec$budg_infl, label=round(mean_budg_by_dec$budg_infl, digits=2), pos=3, cex=0.8, col="red")

mean_prof_plot <- barplot(mean_prof_by_dec$profit_infl, main="Average Profit By Decade", names.arg=mean_prof_by_dec$decade, xlab = "Decade", ylab="Avg Profit")
text(x=mean_prof_plot, y=mean_prof_by_dec$profit_infl, label=round(mean_prof_by_dec$profit_infl, digits=2), pos=3, cex=0.8, col="red")
```
By grouping each of these columns by decade since 1910, the most immediately evident trend is the growth in number of films made in the past 100 years. Despite the growth in movie count since 1980, the average profit and average user rating score has noticeably decreased over the past 40 years - a clear sign that an influx in quantity is not necessarily indicative of a rise in quality. 

# Do films with greater production budgets experience greater profit margins?
```{r}
profit_budget <- lm(m$profit ~ m$budget)
profit_budget_infl <- lm(m$profit_infl ~ m$budg_infl)
summary(profit_budget)
summary(profit_budget_infl)
```
```{r out.width="50%", out.height="50%"}
plot(m$budget, m$profit)
abline(profit_budget, col="red")
title("Relationship Between Budget and Profit")

plot(m$budg_infl, m$profit_infl)
abline(profit_budget_infl, col="red")
title("Relationship Between Inflated Budget and Inflated Profit")
```
To examine the relationship between profit and budget, we created two separate linear regression models: one accounting for inflation and one that does not account for inflation. Profit and budget have a slight positive correlation in both models, but the model without inflation adjustment had a slightly better R-Squared value of **0.3842**.


## Does the budget have an effect on the profit margins, and user ratings of the movie?
```{r}
rating_budget <- lm(m$avg.vote ~ m$budget)
rating_budget_infl <- lm(m$avg.vote  ~ m$budg_infl)
summary(rating_budget)
summary(rating_budget_infl)
```
```{r out.width="50%", out.height="50%"}
plot(m$budget, m$profit)
abline(profit_budget, col="red")
title("Relationship Between Budget and Profit")

plot(m$budg_infl, m$profit_infl)
abline(profit_budget_infl, col="red")
title("Relationship Between Inflated Budget and Inflated Profit")
```
We constructed two more simple linear regression models to examine the relationship between budget, inflated budget, and average user rating score. Much like the previous models, there is a slight positive relationship, but the R squared values in both models indicate that budget alone is not the most effective predictor of rating performance.

## What is the best way to predict the profit of a movie?
To approach best assessing profit, Multiple Linear Regression is chosen as the best analytical method. Models will be assessed using Stepwise regression using both BIC and AIC. Profit will be assessed for movies created in the 21st century as profit seems to level out based on the time series graphs. Dataset will split into two and evaluated separately to account for movies where profit is positive or negative.
A final model will be chosen, if possible, based on above criteria. 
### Data subset and Transformation
```{r}
#removing metascore instead of rows with NA's because of the number of NA's that cause problems in the models.
USAMOVIES <- select(USAMOVIES,-c("metascore"))
#merging dataset to include ratings dataset to incorporate more information that might explain profit
names(ratings)<-str_replace_all(names(ratings), c("_" = "." , "," = "" ))
ratings<- select(ratings,c("imdb.title.id","mean.vote","median.vote","males.allages.avg.vote","males.allages.votes","females.allages.avg.vote","females.allages.votes","top1000.voters.rating","top1000.voters.votes","us.voters.rating","us.voters.votes","non.us.voters.rating","non.us.voters.votes"))
USAMOVIESratings <- merge(USAMOVIES,ratings,by="imdb.title.id")
USAMOVIESratings<- USAMOVIESratings[!is.na(USAMOVIESratings$reviews.from.users),]
USAMOVIESratings<- USAMOVIESratings[!is.na(USAMOVIESratings$reviews.from.critics),]
USAMOVIES1 <- USAMOVIESratings

#subsetting dataset to only account for movies from 2000 on.
year21<-subset(USAMOVIESratings, year>=2000)
#subsetting dataset to separate movies where profit as positive or negative
posprof<- subset(year21, year21$profit.infl>0)
negprof<- subset(year21, year21$profit.infl<=0)
```
Because of the skewedness of the profit variable, profit has been transformed to account for the constant variance assumption in linear regression. Log transformation was orginally not taken into account but corrected upon realizing skewedness.
```{r}
logprofitpos<- log(posprof$profit.infl)
#removal of na values from the negative dataset to account for errors in model.
negprof<- negprof[!is.na(negprof$reviews.from.users),]
negprof<- negprof[!is.na(negprof$reviews.from.critics),]
#Using a cubed root transformation because log and sqrt transformation were not powerful enough due to the severe skewedness of profit.
absprofit<- abs(negprof$profit.infl)
logprofitneg<- (absprofit)^(1/3)
```

### Full Model

Creation of a Full Model from both the positive and negative datasets.
```{r}
fullmodpos<- lm(logprofitpos~ year+ duration +avg.vote+votes+reviews.from.users+reviews.from.critics+IsItHorror+IsItRomance+IsItAction+IsItComedy+IsItDrama+mean.vote+median.vote+males.allages.avg.vote+males.allages.votes+females.allages.avg.vote+females.allages.votes+top1000.voters.rating+top1000.voters.votes+us.voters.rating+us.voters.votes+non.us.voters.rating+non.us.voters.votes, data=posprof)
summary(fullmodpos)

fullmodneg<- lm(logprofitneg~ year+ duration +avg.vote+votes+reviews.from.users+reviews.from.critics+IsItHorror+IsItRomance+IsItAction+IsItComedy+IsItDrama+mean.vote+median.vote+males.allages.avg.vote+males.allages.votes+females.allages.avg.vote+females.allages.votes+top1000.voters.rating+top1000.voters.votes+us.voters.rating+us.voters.votes+non.us.voters.rating+non.us.voters.votes, data=negprof)
summary(fullmodneg)
```

Using stepwise regression for model evaluation. Initially used Backwards and Forwards but appeared redundant and stepwise accounted better than both.

### AIC
```{r}
  StepAICpos <- step(fullmodpos,direction="both", data=posprof)
  StepAICmodpos<-lm(logprofitpos ~ year + duration + avg.vote + votes + reviews.from.users + 
                      reviews.from.critics + IsItRomance + IsItAction + IsItComedy + 
                      IsItDrama + males.allages.avg.vote + females.allages.avg.vote + 
                      females.allages.votes + top1000.voters.rating + top1000.voters.votes + 
                      us.voters.rating + non.us.voters.votes,data=posprof)
  summary(StepAICmodpos)
  
```
Output for positive profit dataset using AIC:
    #AIC= 929.51
    #R^2= 0.597  
```{r}
  StepAICneg <- step(fullmodneg,direction="both", data=negprof)
  StepAICmodneg<-lm(logprofitneg ~ year + duration + reviews.from.critics + IsItHorror + 
                      IsItAction + IsItComedy + IsItDrama + mean.vote + males.allages.avg.vote + 
                      males.allages.votes + females.allages.avg.vote + females.allages.votes + 
                      top1000.voters.votes + us.voters.rating + us.voters.votes + 
                      non.us.voters.votes,data=negprof)
  summary(StepAICmodneg)
```
Output for negative profit dataset using AIC:
  #AIC= 12824.73
  #R^2= 0.3681   
  
### BIC  
Creating variable for BIC evaluation
```{r}  
  npos <- length(fullmodpos$residuals)
  nneg <- length(fullmodneg$residuals)
```


```{r}
  StepBIC <- step(fullmodpos,direction="both", data=posprof, k=log(npos))
  StepBICmodelpos<-lm(logprofitpos ~ year + duration + avg.vote + votes + reviews.from.users + 
                         reviews.from.critics + IsItComedy + IsItDrama + males.allages.avg.vote + 
                         females.allages.avg.vote + females.allages.votes + top1000.voters.votes + 
                         us.voters.rating + non.us.voters.votes,data=posprof)
  summary(StepBICmodelpos)
```
Output for positive profit dataset using BIC:
    #BIC= 1022.99
    #R^2= 0.5955   
```{r}
  StepBIC <- step(fullmodneg,direction="both", data=negprof, k=log(nneg))
  StepBICmodelneg<-lm(logprofitneg ~ duration + reviews.from.critics + IsItHorror + 
                        IsItAction + mean.vote + males.allages.avg.vote + males.allages.votes + 
                        females.allages.avg.vote + females.allages.votes + top1000.voters.votes + 
                        us.voters.rating + us.voters.votes + non.us.voters.votes,data=negprof)
  summary(StepBICmodelneg)
```
Output for negative profit dataset using BIC:
  #BIC= 12905.24
  #R^2= 0.3643  
### Final Model
  The best assessment of profit is StepBICmodpos where BIC= 1022.99 and R^2= 0.5955. The final model has 3 less variables then its counter part using BIC. The difference between the two models R^2 is less than 2% difference, therefore the simpler model was chosen. With this model all variables are significant in the model.
  The variables in the final model include: year,duration,avg.vote,votes,reviews.from.users,reviews.from.critics,IsItComedy,IsItDrama,males.allages.avg.vote,females.allages.avg.vote,females.allages.votes,top1000.voters.votes,us.voters.rating,non.us.voters.votes. 
  A R^2 of 0.5955 that the model does an adequate job of understand why a movie is profitable. However, there is room for improvement outside the scope of the dataset. 
  Even though both model for the negative dataset were significant, based on the R^2 are adequate to explain why a movie did not make money.
  Overall through MLR and separating the dataset, it can be explained why a movie from 2000-2020 is able to make money. However, it cannot explain why a movie would not make money.

## Are production companies a significant variable when discussing the success of a movie?

### Data Exploration

```{r}
summary(USAMOVIESratings[, c("production.company","profit", "profit.infl", "mean.vote")])

# Finding the production companies with the most movies
prod_comp_groups <- count(group_by(USAMOVIESratings,production.company))
prod_co_groups_num_movies <- arrange(prod_comp_groups,-n)
top_companies_list <- prod_co_groups_num_movies[prod_co_groups_num_movies$n > 50,]
movies_by_top_companies <- USAMOVIESratings[USAMOVIESratings$production.company %in% top_companies_list$production.company,]

summary(movies_by_top_companies[,c("production.company", "profit", "profit.infl", "mean.vote")])
summary(prod_co_groups_num_movies)
```

```{r}
attach(movies_by_top_companies)

# Testing the significance of each production company compared to profit 
top.companies.profit.model <- lm(profit~production.company)
summary(top.companies.profit.model)

# Collecting the top 10 production companies that were significant to Profit.
significant_production_cos_profit <- subset(movies_by_top_companies,summary(top.companies.profit.model)$coeff[-1,4] < 0.05)
num_sig_comps_profit <- count(significant_production_cos_profit, vars = production.company)
ordered_run_profit <- num_sig_comps_profit[order(-num_sig_comps_profit$n),]
head(ordered_run_profit, n = 10)

# Testing the significance of each production company compared to profit with inflation
top.companies.profit.inlf.model <- lm(profit.infl~production.company)
summary(top.companies.profit.infl.model)

# Collecting the top 10 production companies that were significant to Profit with inflation.
significant_production_cos_profit_infl <- subset(movies_by_top_companies,summary(top.companies.profit.infl.model)$coeff[-1,4] < 0.05)
num_sig_comps_profit_infl <- count(significant_production_cos_profit_infl, vars = production.company)
ordered_run_profit_infl <- num_sig_comps_profit_infl[order(-num_sig_comps_profit_infl$n),]
head(ordered_run_profit_infl, n = 10)
```

# Ratings Modeling

```{r}
# Testing the significance of each production company compared to ratings

top.companies.ratings.model <- lm(avg.vote ~ production.company)
summary(top.companies.ratings.model)

# Collecting the top 10 production companies that were significant to Ratings.
significant_production_cos_ratings <- subset(movies_by_top_companies,summary(top.companies.ratings.model)$coeff[-1,4] < 0.05)
num_sig_comps_profit_ratings <- count(significant_production_cos_ratings, vars = production.company)
ordered_run_profit_ratings <- num_sig_comps_profit_ratings[order(-num_sig_comps_profit_ratings$n),]
head(ordered_run_profit_ratings, n = 10)
```















## How are ratings affected by movie genre, movie duration, gender?

### Genres
```{r}
attach(USAMOVIES1)
lm_all <- lm(avg.vote ~ IsItHorror + IsItRomance + IsItAction + IsItComedy + IsItDrama)
summary(lm_all)
```
Above, I ran a multiple linear regression that tried to predict the average vote based
on the top five genres I discussed in the previous section. Based on the summary,
the model is significant and all variables are also significant. This tells us that 
the genre of a movie will affect the voting on the quality of the movie. Below, I 
examine using t.tests whether there is a significance between the difference of means
of movies of a specific genre type and movies that do not contain that genre.

## Horror Plots
### T-Tests
```{r}
Horror <- subset(USAMOVIES1, IsItHorror == 1, select = c(avg.vote))
NoHorror <- subset(USAMOVIES1, IsItHorror == 0, select = c(avg.vote))
t.test(Horror,NoHorror)
```
Based on the t-test above, we can see there is a difference between -0.7031330 and 
-0.5389663 with a 95% confidence interval. Because of this interval, there is a significant 
difference in the means of these groups of movies. This shows that horror movies will have 
a lower rating in general compared to other movie types. The histograms below show this spread.
```{r}
p <- ggplot(Horror, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Horror Movie Ratings")
p <- ggplot(NoHorror, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Non-Horror Movie Ratings")
```

## Romance Plots
```{r}
Romance <- subset(USAMOVIES1, IsItRomance == 1, select = c(avg.vote))
NoRomance <- subset(USAMOVIES1, IsItRomance == 0, select = c(avg.vote))
t.test(Romance, NoRomance)
```
Based on the t-test above, we can see there is a difference between -0.07828863 and 
0.03853488 with a 95% confidence interval. Because of this interval, there is not a significant 
difference in the means of these groups of movies due to the fact that our confidence 
interval includes 0. This means that the ratings of a Romance movie and a non-romantic
movie will not have a significant difference in mean based on the genre.The histograms
below show the spread of the data.

```{r}
p <- ggplot(Romance, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Romance Movie Ratings")
            
p <- ggplot(NoRomance, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Non-Romance Movie Ratings")
```


## Action Plots
```{r}
Action <- subset(USAMOVIES1, IsItAction == 1, select = c(avg.vote))
NoAction <- subset(USAMOVIES1, IsItAction == 0, select = c(avg.vote))
t.test(Action,NoAction)
```
Based on the t-test above, we can see there is a difference between -0.17345224 and 
-0.05673903 with a 95% confidence interval. Because of this interval, there is a significant 
difference in the means of these groups of movies. This shows that action movies will have 
a slightly lower rating in general compared to other movie types. The histograms below show this spread.

```{r}
p <- ggplot(Action, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Action Movie Ratings")
            
p <- ggplot(NoAction, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Non-Action Movie Ratings")
```

## Comedy Plot
```{r}
Comedy <- subset(USAMOVIES1, IsItComedy == 1, select = c(avg.vote))
NoComedy <- subset(USAMOVIES1, IsItComedy == 0, select = c(avg.vote))
t.test(Comedy,NoComedy)
```
Based on the t-test above, we can see there is a difference between -0.3428530 and 
-0.2429231 with a 95% confidence interval. Because of this interval, there is a significant 
difference in the means of these groups of movies. This shows that comedy movies will have 
a lower rating in general compared to other movie types. The histograms below show this spread.

```{r}
p <- ggplot(Comedy, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Comedy Movie Ratings")
            
p <- ggplot(NoComedy, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Non-Comedy Movie Ratings")
```

## Drama Plot
```{r}
Drama <- subset(USAMOVIES1, IsItDrama == 1, select = c(avg.vote))
NoDrama <- subset(USAMOVIES1, IsItDrama == 0, select = c(avg.vote))
t.test(Drama,NoDrama)
```
Based on the t-test above, we can see there is a difference between 0.4930031 and 
0.5904080 with a 95% confidence interval. Because of this interval, there is a significant 
difference in the means of these groups of movies. This shows that dramatic movies will have 
a higher rating in general compared to other movie types. The histograms below show this spread.

```{r}
p <- ggplot(Drama, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Drama Movie Ratings")
            
p <- ggplot(NoDrama, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Non-Drama Movie Ratings")
```


## How are Ratings Affected by Gender?
We shift gears in this section from genres into whether gender plays a role in the 
average rating of a movie. The t-test below checks to see if their is a significant 
difference in the means of the female's average vote and the male's average vote. 
We can see from the test that females vote more positivitely for movies on average
in a range from 0.1812546 to 0.2513334 on a 95% percent confidence interval. 
```{r}
t.test(females.allages.avg.vote,males.allages.avg.vote)
```
The histogram below shows the spread of the two genders' votes. The male vote is indicated 
with blue bars. The female vote is represented using pink bars. The overlap in the votes
is indicated using the purple bars. We can see from the histogram the confirmation of the 
observation above that females votes more positively for movies on average than males.
```{r}
hgf <- hist(females.allages.avg.vote,plot = FALSE)
hgm <- hist(males.allages.avg.vote,plot = FALSE)

plot(hgf,col = rgb(1,0,0,.2), main = "Spread of Female and Male Vote")
plot(hgm,col = rgb(0,0,1,.2), add=TRUE)
```


In the next two sections, I wanted to combine the ideas from the two sections above.
Is there a significant difference in average voting for a movie based on genres and gender.
I chose to look specifically at two stereotypical movie and gender combinations to see
if the stereotype would hold true when looking at the data. These combinations were
women and romance movies and men and action movies. Below, I will examine if there is 
a significant difference in the data when looking at males and females votes specifically in the romance
and action movie genres.
## Do Females Actually Like Romantic Movies More than Males?
```{r}
Romance1 <- subset(USAMOVIES1, IsItRomance == 1, select = c(avg.vote,males.allages.avg.vote,females.allages.avg.vote))
t.test(Romance1$females.allages.avg.vote,Romance1$males.allages.avg.vote)
```
We can see from the t-test above that women will vote between 0.1912656 and 0.3371316
points higher than males on movies in the romance category in a 95% confidence interval.
Below you can see that spread in a box plot:
```{r}
boxplot(Romance1$females.allages.avg.vote,Romance1$males.allages.avg.vote,
names = c("Female","Male"),
col = c("pink","blue"))
```

## Do Males Actually Like Action Movies More than Females?
```{r}
Action1 <- subset(USAMOVIES1, IsItAction == 1, select = c(avg.vote,males.allages.avg.vote,females.allages.avg.vote))
t.test(Action1$females.allages.avg.vote,Action1$males.allages.avg.vote)
```
We can see from the t-test above that women will vote between 0.1003357 and 0.2405763
points higher than males on movies in the female category in a 95% confidence interval.
This contradicts the stereotype that males like action movies more than females.
Below you can see that spread in a box plot:
```{r}
boxplot(Action1$females.allages.avg.vote,Action1$males.allages.avg.vote,
names = c("Female","Male"),
col = c("pink","blue"))
```

## How Does Movie Duration Affect a Movie's Success?
In the section below, we will look to see if there is any correlation between the duration
of movies and the average votes for the movie.
### Initial Model
```{r}
duration_lm <- lm(avg.vote~ duration)
summary(duration_lm)
plot(duration,avg.vote)
abline(duration_lm, col = 'red')
```

### Model after taking out some outlier values
```{r}
res_std <- rstandard(duration_lm)
USAMOVIES1$res_std = res_std
USAMOVIES2 <- USAMOVIES1[USAMOVIES1$res_std <= 2,]
USAMOVIES3 <- USAMOVIES2[USAMOVIES1$res_std >= -2,]
duration_lm2 <- lm(USAMOVIES3$avg.vote~ USAMOVIES3$duration)
summary(duration_lm2)
plot(USAMOVIES3$duration,USAMOVIES3$avg.vote)
abline(duration_lm2, col= 'red')
```

We can see from the summary above that the model and the variable for duration is 
significant when predicting the average vote.




