---
title: "ST_541_FinalProject"
author: "Collins Mills, Mary Strickland, Carter Lockwood, Steven Johnson"
date: "December 9th 2020"
output: pdf_document
---
# Data Insert and Cleaning
```{r setup, include=FALSE}
library(readr)
library(leaps)
library("stringr")
library(ggplot2)
library(car)
library(dplyr)

IMDb_movies <- read.csv(file = "IMDb movies.csv", header=TRUE)
ratings <- read.csv(file = "IMDb ratings.csv", header=TRUE)
cpi_rates <- read.csv(file = "cpi_table.csv", header=TRUE)
```
## Renaming and inserting columns
```{r}
names(ratings)<-str_replace_all(names(ratings), c("_" = "." , "," = "" ))
names(IMDb_movies)<-str_replace_all(names(IMDb_movies), c("_" = "." , "," = "" ))
#rename 'worlwide' to 'worldwide'
colnames(IMDb_movies)[19] <- "worldwide.gross.income"
#Inserting dummy variables to account for Movie Genres originally combined into one column.
IMDb_movies["IsItHorror"]<- 0
IMDb_movies["IsItRomance"] <- 0
IMDb_movies["IsItAction"] <- 0
IMDb_movies["IsItComedy"] <- 0
IMDb_movies["IsItDrama"] <- 0
IMDb_movies$IsItHorror[grep("Horror",IMDb_movies$genre)] <- 1
IMDb_movies$IsItRomance[grep("Romance",IMDb_movies$genre)] <- 1
IMDb_movies$IsItAction[grep("Action",IMDb_movies$genre)] <- 1
IMDb_movies$IsItComedy[grep("Comedy",IMDb_movies$genre)] <- 1
IMDb_movies$IsItDrama[grep("Drama",IMDb_movies$genre)] <- 1
IMDb_movies$IsItHorror<-as.factor(IMDb_movies$IsItHorror)
IMDb_movies$IsItRomance<-as.factor(IMDb_movies$IsItRomance)
IMDb_movies$IsItAction<-as.factor(IMDb_movies$IsItAction)
IMDb_movies$IsItComedy<-as.factor(IMDb_movies$IsItComedy)
IMDb_movies$IsItDrama<-as.factor(IMDb_movies$IsItDrama)
```

## Data Subsetting

```{r}
#To account for NA records and extracting only records where the movie was made in the USA. 
#This decision was made to focus analysis and minimize the dataset. 
budget <-IMDb_movies[!is.na(IMDb_movies$budget) & !(IMDb_movies$budget==""),]
usamovies<-budget[!is.na(budget$usa.gross.income) & !(budget$usa.gross.income == ""),]
usamovies1<-usamovies[!is.na(usamovies$country) & !(usamovies$country==""),]
usamovies2<-usamovies1[grep("USA",usamovies1$country),]
#Inserting a currency column to account for only records where profit was measure in dollars.
# This eliminated the need to convert currency and created consistency in the dataset.
usamovies2$Currency <- substr(usamovies2$budget, start = 1, stop = 3)
for(value in 1:6605){
  if(substr(usamovies2$Currency[value], start = 1, stop = 1)=="$"){
    usamovies2$Currency[value] = "USD"
  }
}
USAMOVIES<-subset(usamovies2, usamovies2$Currency== "USD")
#Recreating budget and gross income columns to remove "$" and change type to numeric
USAMOVIES$budget <- substr(USAMOVIES$budget, start = 2, stop = 15)
USAMOVIES$usa.gross.income <- substr(USAMOVIES$usa.gross.income, start = 2, stop = 15)
USAMOVIES$worldwide.gross.income <- substr(USAMOVIES$worldwide.gross.income, start = 2, stop = 15)
USAMOVIES$budget <- as.numeric(USAMOVIES$budget)
USAMOVIES$usa.gross.income <- as.numeric(USAMOVIES$usa.gross.income)
USAMOVIES$worldwide.gross.income <- as.numeric(USAMOVIES$worldwide.gross.income)
#Creating of profit column under the assumption that all budget was used and world wide gross income is the best representation of money earned.
USAMOVIES$profit <-(USAMOVIES$worldwide.gross.income- USAMOVIES$budget)

#make year column into the index for cpi_rates dataframe
row.names(cpi_rates) <- cpi_rates$Year
cpi_rates[1] <- NULL
cpi_2020 <-  cpi_rates["2020", 1]


#Formula for inflated values: (original value * 2020 Consumer Price Index Value) / Consumer Price Index From Year of Release
for(val in 1:nrow(USAMOVIES)){
  USAMOVIES[val, "budg.infl"] <- (USAMOVIES[val, "budget"] * cpi_2020) / cpi_rates[toString(USAMOVIES[val, "year"]), "Avg"]
  USAMOVIES[val, "worldwide.gross.infl"] <- (USAMOVIES[val, "worldwide.gross.income"] * cpi_2020) / cpi_rates[toString(USAMOVIES[val, "year"]), "Avg"]
  USAMOVIES[val, "usa.gross.infl"] <- (USAMOVIES[val, "usa.gross.income"] * cpi_2020) / cpi_rates[toString(USAMOVIES[val, "year"]), "Avg"]
  USAMOVIES[val, "profit.infl"] <- (USAMOVIES[val, "profit"] * cpi_2020) / cpi_rates[toString(USAMOVIES[val, "year"]), "Avg"]
}

#Creating the data set with movie data and ratings data 
#removing metascore instead of rows with NA's because of the number of NA's that cause problems in the models.
USAMOVIES <- select(USAMOVIES,-c("metascore"))
#merging dataset to include ratings dataset to incorporate more information that might explain profit
names(ratings)<-str_replace_all(names(ratings), c("_" = "." , "," = "" ))
ratings<- select(ratings,c("imdb.title.id","mean.vote","median.vote","males.allages.avg.vote","males.allages.votes","females.allages.avg.vote","females.allages.votes","top1000.voters.rating","top1000.voters.votes","us.voters.rating","us.voters.votes","non.us.voters.rating","non.us.voters.votes"))
USAMOVIESratings <- merge(USAMOVIES,ratings,by="imdb.title.id")
USAMOVIESratings<- USAMOVIESratings[!is.na(USAMOVIESratings$reviews.from.users),]
USAMOVIESratings<- USAMOVIESratings[!is.na(USAMOVIESratings$reviews.from.critics),]
USAMOVIES1 <- USAMOVIESratings

summary(USAMOVIES)
```
# Successful Movies
## Getting familiar with the data: Examining Count, Budget, Profit, and Score Over The Years
```{r}
decades <- c(1910, 1920, 1930, 1940, 1950, 1960, 1970, 1980, 1990, 2000, 2010, 2020)
USAMOVIES$decade <- decades[findInterval(USAMOVIES$year, decades)]
mean_budg_by_dec <- aggregate(budg.infl ~ decade, data = USAMOVIES, mean)
mean_prof_by_dec <- aggregate(profit.infl ~ decade, data = USAMOVIES, mean)
mean_score_by_dec <- aggregate(avg.vote ~ decade, data = USAMOVIES, mean)
count_by_decade_plot<- barplot(table(USAMOVIES$decade), main="Movie Count By Decade", xlab="Decade", ylab="Number of Movies")

mean_score_plot <- barplot(mean_score_by_dec$avg.vote, main="Average Score By Decade", names.arg=mean_score_by_dec$decade, xlab = "Decade", ylab="Avg Score")
text(x=mean_score_plot, y=mean_score_by_dec$avg.vote, label=round(mean_score_by_dec$avg.vote, digits=2), pos=3, cex=0.8, col="red")

mean_budg_plot <- barplot(mean_budg_by_dec$budg.infl, main="Average Budget By Decade", names.arg=mean_budg_by_dec$decade, xlab = "Decade", ylab="Avg Budget")
text(x=mean_budg_plot, y=mean_budg_by_dec$budg.infl, label=round(mean_budg_by_dec$budg.infl, digits=2), pos=3, cex=0.8, col="red")

mean_prof_plot <- barplot(mean_prof_by_dec$profit.infl, main="Average Profit By Decade", names.arg=mean_prof_by_dec$decade, xlab = "Decade", ylab="Avg Profit")
text(x=mean_prof_plot, y=mean_prof_by_dec$profit.infl, label=round(mean_prof_by_dec$profit.infl, digits=2), pos=3, cex=0.8, col="red")
```
By grouping each of these columns by decade since 1910, the most immediately evident trend is the growth in number of films made in the past 100 years. Despite the growth in movie count since 1980, the average profit and average user rating score has noticeably decreased over the past 40 years - a clear sign that an influx in quantity is not necessarily indicative of a rise in quality. 

# Do films with greater production budgets experience greater profit margins?
```{r}
profit_budget <- lm(USAMOVIES$profit ~ USAMOVIES$budget)
profit_budget_infl <- lm(USAMOVIES$profit.infl ~ USAMOVIES$budg.infl)
summary(profit_budget)
summary(profit_budget_infl)
```
```{r out.width="50%", out.height="50%"}
plot(USAMOVIES$budget, USAMOVIES$profit)
abline(profit_budget, col="red")
title("Relationship Between Budget and Profit")

plot(USAMOVIES$budg.infl, USAMOVIES$profit.infl)
abline(profit_budget_infl, col="red")
title("Relationship Between Inflated Budget and Inflated Profit")
```
To examine the relationship between profit and budget, we created two separate linear regression models: one accounting for inflation and one that does not account for inflation. Profit and budget have a slight positive correlation in both models, but the model without inflation adjustment had a slightly better R-Squared value of **0.3842**.


## Does the budget have an effect on the profit margins, and user ratings of the movie?
```{r}
rating_budget <- lm(USAMOVIES$avg.vote ~ USAMOVIES$budget)
rating_budget_infl <- lm(USAMOVIES$avg.vote  ~ USAMOVIES$budg.infl)
summary(rating_budget)
summary(rating_budget_infl)
```
```{r out.width="50%", out.height="50%"}
plot(USAMOVIES$budget, USAMOVIES$profit)
abline(profit_budget, col="red")
title("Relationship Between Budget and Profit")

plot(USAMOVIES$budg.infl, USAMOVIES$profit.infl)
abline(profit_budget_infl, col="red")
title("Relationship Between Inflated Budget and Inflated Profit")
```
We constructed two more simple linear regression models to examine the relationship between budget, inflated budget, and average user rating score. Much like the previous models, there is a slight positive relationship, but the R squared values in both models indicate that budget alone is not the most effective predictor of rating performance.

# Time Series Analysis of Profit and Profit (with Inflation)

Since profit is the response variable we chose to predict, it is important to fully understand that data. The following section will explore more of the profit and profit with inflation variables of the data set.

## Setup for Time Series Analysis
```{r}

# Creating a time series analysis
MOVIES.TIME.SERIES <- USAMOVIESratings[c("year","date.published","profit.infl","profit")]
# Putting the movies in order by the date they were published
MOVIES.TIME.SERIES <- MOVIES.TIME.SERIES[order(MOVIES.TIME.SERIES$date.published),]                             
attach(MOVIES.TIME.SERIES)

# Found inconsistencies in the year and date.published so create a new column to extract the year from date.published
MOVIES.TIME.SERIES$year.published<- as.integer(substring(MOVIES.TIME.SERIES$date.published,1,4))
head(MOVIES.TIME.SERIES$year.published)

# Group movies by year and investigate that time series
years <- group_by(MOVIES.TIME.SERIES,year.published)

```

In order to get a clear picture of how profit changed over time we aggregated the data set by grouping all movies for the year and taking the average of the profit for each year. The following graphs show the relationship of profit and profit for inflation over the years.

## Time Series by year
```{r}
# Graphing the time series of average profit by year with and without inflation.

# Create a set of data that only includes the year and the average profit for that year.
avg.profit.by.year <- summarize(years,avg.profit = mean(profit))

ggplot(data = avg.profit.by.year, aes(x = year.published, y = avg.profit)) +
  geom_line(stat = "identity") +
  labs(title = "Avgerage Profit gained per Year",
       subtitle = "1921 - 2020",x = "Date", y = "Profit")


# Create a set of data that only includes the year and the average profit for that year with inflation.
avg.profit.infl.by.year <- summarize(years,avg.profit.infl = mean(profit.infl))

# Graphing the time series of average profit with inflation by year.

ggplot(data = avg.profit.infl.by.year, aes(x = year.published, y = avg.profit.infl)) +
  geom_line(stat = "identity") +
  labs(title = "Avgerage Profit (with Inflation) gained per Year",
       subtitle = "1921 - 2020",
       x = "Date", y = "Profit")

```

From these graphs we see a lot of variation at the beginning of our data. We also see that profit with inflation begins to level off around the year 1980. To look a little further, we grouped the data again by decade.

## Time Series by Decade
```{r}
# Created a new column that created the decade the movie was created.
MOVIES.TIME.SERIES$decade.published <- as.integer(substring(MOVIES.TIME.SERIES$date.published,1,3))*10

# Grouped by Decade
decade <- group_by(MOVIES.TIME.SERIES,decade.published)

# Calculated the average profit by decade
avg.profit.by.decade <- summarize(decade,avg.profit = mean(profit))

#Graphed the result Profit by decade
ggplot(data = avg.profit.by.decade, aes(x = decade.published, y = avg.profit)) +
  geom_line(stat = "identity") +
  xlim(1920,2020) +
  labs(title = "Avgerage Profit gained over Decades",
       subtitle = "1920 - 2020",
       x = "Date", y = "Profit")

# Calculated the average profit with inflation by decade
avg.profit.infl.by.decade <- summarize(decade,avg.profit.infl = mean(profit.infl))

# Graphed the result Profit with inflation
ggplot(data = avg.profit.infl.by.decade, aes(x = decade.published, y = avg.profit.infl)) +
  geom_line(stat = "identity") +
  xlim(1920,2020) +
  labs(title = "Avgerage Profit with Inflation gained over Decades",
       subtitle = "1920 - 2020",
       x = "Date", y = "Profit")
```
Looking at the decade graphs, profit without inflation has a generally increasing trend which makes intuitive sense because prices were rising over the century. However, the profit with inflation still has a large increase around 1930 and 1940 before it begins to normalize in the second half of the century. We looked into this a little further by creating a faucet wrap shown belown.

```{r}
# Distribution of movies by profit over every decade
ggplot(data=decade, aes(x = profit.infl)) +
  geom_histogram() + 
  ylim(0,10) +
  facet_wrap(~decade.published)
```

From this display of graphs we see that there was a movie in both the 1930s and the 1940s that produced a higher profit than any movie we currently have available. This explains the massive spike in average over those decades. 

## Correlation between the years
```{r}
#Checking for correlation between any given date and the one before.
plot(profit.infl[1:6529], profit.infl[2:6530], pch=20, col=4,
     main="Profit for Individual Movies by Date Published", xlab="Profit(t-1)",
     ylab = "Profit(t)")
text(x=6, y=45, col=2, cex=1.5,
     labels=paste("Corr =", round(cor(profit.infl[1:6529],
                                      profit.infl[2:6530]),2)))

# Autocorrelation between each date
print(acf(profit.infl, main = "Correlation of Profit from one year to the year before"))

```
When looking into correlation between the years, we find that one year does not have a lot of effect on the year before it or after it. However, when running the auto correlation, we do find some year gaps that are labeled significant although their correlations are very low. This begins with about every seven years and happens again around year ten and thirteen.

Now that we know a little more about one of our response variables, we can discuss how to find the correct predictors.

# What is the best way to predict the profit of a movie?
To approach best assessing profit, Multiple Linear Regression is chosen as the best analytical method. Models will be assessed using Stepwise regression using both BIC and AIC. Profit will be assessed for movies created in the 21st century as profit seems to level out based on the time series graphs. Dataset will split into two and evaluated separately to account for movies where profit is positive or negative.
A final model will be chosen, if possible, based on above criteria. 

## Data subset and Transformation
```{r}
#subsetting dataset to only account for movies from 2000 on.
year21<-subset(USAMOVIESratings, year>=2000)
#subsetting dataset to separate movies where profit as positive or negative
posprof<- subset(year21, year21$profit.infl>0)
negprof<- subset(year21, year21$profit.infl<=0)
```
Because of the skewedness of the profit variable, profit has been transformed to account for the constant variance assumption in linear regression. Log transformation was orginally not taken into account but corrected upon realizing skewedness.
```{r}
logprofitpos<- log(posprof$profit.infl)
#removal of na values from the negative dataset to account for errors in model.
negprof<- negprof[!is.na(negprof$reviews.from.users),]
negprof<- negprof[!is.na(negprof$reviews.from.critics),]
#Using a cubed root transformation because log and sqrt transformation were not powerful enough due to the severe skewedness of profit.
absprofit<- abs(negprof$profit.infl)
logprofitneg<- (absprofit)^(1/3)
```

## Full Model

Creation of a Full Model from both the positive and negative datasets.
```{r}
fullmodpos<- lm(logprofitpos~ year+ duration +avg.vote+votes+reviews.from.users+reviews.from.critics+IsItHorror+IsItRomance+IsItAction+IsItComedy+IsItDrama+mean.vote+median.vote+males.allages.avg.vote+males.allages.votes+females.allages.avg.vote+females.allages.votes+top1000.voters.rating+top1000.voters.votes+us.voters.rating+us.voters.votes+non.us.voters.rating+non.us.voters.votes, data=posprof)
summary(fullmodpos)

fullmodneg<- lm(logprofitneg~ year+ duration +avg.vote+votes+reviews.from.users+reviews.from.critics+IsItHorror+IsItRomance+IsItAction+IsItComedy+IsItDrama+mean.vote+median.vote+males.allages.avg.vote+males.allages.votes+females.allages.avg.vote+females.allages.votes+top1000.voters.rating+top1000.voters.votes+us.voters.rating+us.voters.votes+non.us.voters.rating+non.us.voters.votes, data=negprof)
summary(fullmodneg)
```

Using stepwise regression for model evaluation. Initially used Backwards and Forwards but appeared redundant and stepwise accounted better than both.

## AIC
```{r}
  StepAICpos <- step(fullmodpos,direction="both", data=posprof)
  StepAICmodpos<-lm(logprofitpos ~ year + duration + avg.vote + votes + reviews.from.users + 
                      reviews.from.critics + IsItRomance + IsItAction + IsItComedy + 
                      IsItDrama + males.allages.avg.vote + females.allages.avg.vote + 
                      females.allages.votes + top1000.voters.rating + top1000.voters.votes + 
                      us.voters.rating + non.us.voters.votes,data=posprof)
  summary(StepAICmodpos)
  
```
Output for positive profit dataset using AIC:
    #AIC= 929.51
    #R^2= 0.597  
```{r}
  StepAICneg <- step(fullmodneg,direction="both", data=negprof)
  StepAICmodneg<-lm(logprofitneg ~ year + duration + reviews.from.critics + IsItHorror + 
                      IsItAction + IsItComedy + IsItDrama + mean.vote + males.allages.avg.vote + 
                      males.allages.votes + females.allages.avg.vote + females.allages.votes + 
                      top1000.voters.votes + us.voters.rating + us.voters.votes + 
                      non.us.voters.votes,data=negprof)
  summary(StepAICmodneg)
```
Output for negative profit dataset using AIC:
  #AIC= 12824.73
  #R^2= 0.3681   
  
## BIC  
Creating variable for BIC evaluation
```{r}  
  npos <- length(fullmodpos$residuals)
  nneg <- length(fullmodneg$residuals)
```


```{r}
  StepBIC <- step(fullmodpos,direction="both", data=posprof, k=log(npos))
  StepBICmodelpos<-lm(logprofitpos ~ year + duration + avg.vote + votes + reviews.from.users + 
                         reviews.from.critics + IsItComedy + IsItDrama + males.allages.avg.vote + 
                         females.allages.avg.vote + females.allages.votes + top1000.voters.votes + 
                         us.voters.rating + non.us.voters.votes,data=posprof)
  summary(StepBICmodelpos)
```
Output for positive profit dataset using BIC:
    #BIC= 1022.99
    #R^2= 0.5955   
```{r}
  StepBIC <- step(fullmodneg,direction="both", data=negprof, k=log(nneg))
  StepBICmodelneg<-lm(logprofitneg ~ duration + reviews.from.critics + IsItHorror + 
                        IsItAction + mean.vote + males.allages.avg.vote + males.allages.votes + 
                        females.allages.avg.vote + females.allages.votes + top1000.voters.votes + 
                        us.voters.rating + us.voters.votes + non.us.voters.votes,data=negprof)
  summary(StepBICmodelneg)
```
Output for negative profit dataset using BIC:
  #BIC= 12905.24
  #R^2= 0.3643  
  
## Final Model

  The best assessment of profit is StepBICmodpos where BIC= 1022.99 and R^2= 0.5955. The final model has 3 less variables then its counter part using BIC. The difference between the two models R^2 is less than 2% difference, therefore the simpler model was chosen. With this model all variables are significant in the model.
  The variables in the final model include: year,duration,avg.vote,votes,reviews.from.users,reviews.from.critics,IsItComedy,IsItDrama,males.allages.avg.vote,females.allages.avg.vote,females.allages.votes,top1000.voters.votes,us.voters.rating,non.us.voters.votes. 
  A R^2 of 0.5955 that the model does an adequate job of understand why a movie is profitable. However, there is room for improvement outside the scope of the dataset. 
  Even though both model for the negative dataset were significant, based on the R^2 are adequate to explain why a movie did not make money.
  Overall through MLR and separating the dataset, it can be explained why a movie from 2000-2020 is able to make money. However, it cannot explain why a movie would not make money.

# Are production companies a significant variable when discussing the success of a movie?

## Data Exploration

```{r}
summary(USAMOVIESratings[, c("production.company","profit", "profit.infl", "mean.vote")])


production_companies <- lm(USAMOVIESratings$profit ~ USAMOVIESratings$production.company)
summary(production_companies)
```

From this model it is evident that there are too many independent variables. To 
help subset the data it is helpful to select the top production companies in the data.

The following code will group the data by production company and count the number of 
movies from each production company.

```{r}
# Finding the production companies with the most movies
prod_comp_groups <- count(group_by(USAMOVIESratings,production.company))
prod_co_groups_num_movies <- arrange(prod_comp_groups,-n)
prod_co_groups_num_movies
summary(prod_co_groups_num_movies)
```

The resulting summary shows that 75% of the production companies have only produced one movie. 
The summary also shows that the maximum number of movies produced by a company is 339 movies.
From the ordered groupings we see that there is a good cut off around 50 movies.
Therefore, we subset the data based on the 18 production companies that made at least 
50 movies in the data set.

```{r}
top_companies_list <- prod_co_groups_num_movies[prod_co_groups_num_movies$n > 50,]
movies_by_top_companies <- USAMOVIESratings[USAMOVIESratings$production.company %in% top_companies_list$production.company,]

summary(movies_by_top_companies[,c("production.company", "profit", "profit.infl", "mean.vote")])

```

After subsetting the data and looking at the new summaries it looked good to move forward in the analysis.

## Production Company Profit Modeling

```{r}
attach(movies_by_top_companies)

# Testing the significance of each production company compared to profit 
top.companies.profit.model <- lm(profit~production.company)
summary(top.companies.profit.model)
```

This model shows that there are eight significant production companies. Walt Disney Pictures, Twentieth Century Fox, and Warner Brothers were all positively significant while Fox Searchlight Pictures, Metro-Goldwyn-Mayer (MGM), Miramax, Screen Gems, and TriStar Pictures had a signifiant negative relationship with profit.

```{r}

# Testing the significance of each production company compared to profit with inflation
top.companies.profit.inlf.model <- lm(profit.infl~production.company)
summary(top.companies.profit.inlf.model)

```

With another significant model, production companies continue to have an affect on the profit gained. However, in this model we see that there was some fluctuation of which companies where positively and negatively significant.Paramount Pictures and Universal Pictures joined Walt Disney Pictures, Twentieth Century Fox, and Warner Brothers as positive influences for this model. Fox Searchlight Pictures, Miramax, and Screen Gems remained significantly negative for this model.

## Production Company Ratings Modeling

```{r}
# Testing the significance of each production company compared to ratings

top.companies.ratings.model <- lm(avg.vote ~ production.company)
summary(top.companies.ratings.model)


```
 
According to this model, production companies have a significant impact on the user ratings. There was one less significant company that was in this model compared to the profit models. Warner Brothers, DreamWorks,Fox Searchlight Pictures and Miramax made a positively significant impact. While Dimension Films, New Line Cinema, and Screen Gems had a reverse effect on the average vote. This places Screen Gems as a negative factor for both profit and average vote.

# How are ratings affected by movie genre, movie duration, gender?

## Genres

```{r}
attach(USAMOVIES1)
lm_all <- lm(avg.vote ~ IsItHorror + IsItRomance + IsItAction + IsItComedy + IsItDrama)
summary(lm_all)
```
Above, I ran a multiple linear regression that tried to predict the average vote based
on the top five genres I discussed in the previous section. Based on the summary,
the model is significant and all variables are also significant. This tells us that 
the genre of a movie will affect the voting on the quality of the movie. Below, I 
examine using t.tests whether there is a significance between the difference of means
of movies of a specific genre type and movies that do not contain that genre.

## Horror Plots
### T-Tests
```{r}
Horror <- subset(USAMOVIES1, IsItHorror == 1, select = c(avg.vote))
NoHorror <- subset(USAMOVIES1, IsItHorror == 0, select = c(avg.vote))
t.test(Horror,NoHorror)
```
Based on the t-test above, we can see there is a difference between -0.7031330 and 
-0.5389663 with a 95% confidence interval. Because of this interval, there is a significant 
difference in the means of these groups of movies. This shows that horror movies will have 
a lower rating in general compared to other movie types. The histograms below show this spread.
```{r}
p <- ggplot(Horror, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Horror Movie Ratings")
p <- ggplot(NoHorror, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Non-Horror Movie Ratings")
```

## Romance Plots
```{r}
Romance <- subset(USAMOVIES1, IsItRomance == 1, select = c(avg.vote))
NoRomance <- subset(USAMOVIES1, IsItRomance == 0, select = c(avg.vote))
t.test(Romance, NoRomance)
```
Based on the t-test above, we can see there is a difference between -0.07828863 and 
0.03853488 with a 95% confidence interval. Because of this interval, there is not a significant 
difference in the means of these groups of movies due to the fact that our confidence 
interval includes 0. This means that the ratings of a Romance movie and a non-romantic
movie will not have a significant difference in mean based on the genre.The histograms
below show the spread of the data.

```{r}
p <- ggplot(Romance, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Romance Movie Ratings")
            
p <- ggplot(NoRomance, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Non-Romance Movie Ratings")
```


## Action Plots
```{r}
Action <- subset(USAMOVIES1, IsItAction == 1, select = c(avg.vote))
NoAction <- subset(USAMOVIES1, IsItAction == 0, select = c(avg.vote))
t.test(Action,NoAction)
```
Based on the t-test above, we can see there is a difference between -0.17345224 and 
-0.05673903 with a 95% confidence interval. Because of this interval, there is a significant 
difference in the means of these groups of movies. This shows that action movies will have 
a slightly lower rating in general compared to other movie types. The histograms below show this spread.

```{r}
p <- ggplot(Action, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Action Movie Ratings")
            
p <- ggplot(NoAction, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Non-Action Movie Ratings")
```

## Comedy Plot
```{r}
Comedy <- subset(USAMOVIES1, IsItComedy == 1, select = c(avg.vote))
NoComedy <- subset(USAMOVIES1, IsItComedy == 0, select = c(avg.vote))
t.test(Comedy,NoComedy)
```
Based on the t-test above, we can see there is a difference between -0.3428530 and 
-0.2429231 with a 95% confidence interval. Because of this interval, there is a significant 
difference in the means of these groups of movies. This shows that comedy movies will have 
a lower rating in general compared to other movie types. The histograms below show this spread.

```{r}
p <- ggplot(Comedy, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Comedy Movie Ratings")
            
p <- ggplot(NoComedy, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Non-Comedy Movie Ratings")
```

## Drama Plot
```{r}
Drama <- subset(USAMOVIES1, IsItDrama == 1, select = c(avg.vote))
NoDrama <- subset(USAMOVIES1, IsItDrama == 0, select = c(avg.vote))
t.test(Drama,NoDrama)
```
Based on the t-test above, we can see there is a difference between 0.4930031 and 
0.5904080 with a 95% confidence interval. Because of this interval, there is a significant 
difference in the means of these groups of movies. This shows that dramatic movies will have 
a higher rating in general compared to other movie types. The histograms below show this spread.

```{r}
p <- ggplot(Drama, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Drama Movie Ratings")
            
p <- ggplot(NoDrama, aes(x=avg.vote)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") + scale_x_continuous(breaks=seq(0, 10, 0.5))
 p+ geom_vline(aes(xintercept=mean(avg.vote)),
            color="blue", linetype="dashed", size=1) + ggtitle("Spread of Non-Drama Movie Ratings")
```


## How are Ratings Affected by Gender?
We shift gears in this section from genres into whether gender plays a role in the 
average rating of a movie. The t-test below checks to see if their is a significant 
difference in the means of the female's average vote and the male's average vote. 
We can see from the test that females vote more positivitely for movies on average
in a range from 0.1812546 to 0.2513334 on a 95% percent confidence interval. 
```{r}
t.test(females.allages.avg.vote,males.allages.avg.vote)
```
The histogram below shows the spread of the two genders' votes. The male vote is indicated 
with blue bars. The female vote is represented using pink bars. The overlap in the votes
is indicated using the purple bars. We can see from the histogram the confirmation of the 
observation above that females votes more positively for movies on average than males.
```{r}
hgf <- hist(females.allages.avg.vote,plot = FALSE)
hgm <- hist(males.allages.avg.vote,plot = FALSE)

plot(hgf,col = rgb(1,0,0,.2), main = "Spread of Female and Male Vote")
plot(hgm,col = rgb(0,0,1,.2), add=TRUE)
```


In the next two sections, I wanted to combine the ideas from the two sections above.
Is there a significant difference in average voting for a movie based on genres and gender.
I chose to look specifically at two stereotypical movie and gender combinations to see
if the stereotype would hold true when looking at the data. These combinations were
women and romance movies and men and action movies. Below, I will examine if there is 
a significant difference in the data when looking at males and females votes specifically in the romance
and action movie genres.

## Do Females Actually Like Romantic Movies More than Males?

```{r}
Romance1 <- subset(USAMOVIES1, IsItRomance == 1, select = c(avg.vote,males.allages.avg.vote,females.allages.avg.vote))
t.test(Romance1$females.allages.avg.vote,Romance1$males.allages.avg.vote)
```
We can see from the t-test above that women will vote between 0.1912656 and 0.3371316
points higher than males on movies in the romance category in a 95% confidence interval.
Below you can see that spread in a box plot:
```{r}
boxplot(Romance1$females.allages.avg.vote,Romance1$males.allages.avg.vote,
names = c("Female","Male"),
col = c("pink","blue"))
```

## Do Males Actually Like Action Movies More than Females?

```{r}
Action1 <- subset(USAMOVIES1, IsItAction == 1, select = c(avg.vote,males.allages.avg.vote,females.allages.avg.vote))
t.test(Action1$females.allages.avg.vote,Action1$males.allages.avg.vote)
```
We can see from the t-test above that women will vote between 0.1003357 and 0.2405763
points higher than males on movies in the female category in a 95% confidence interval.
This contradicts the stereotype that males like action movies more than females.
Below you can see that spread in a box plot:
```{r}
boxplot(Action1$females.allages.avg.vote,Action1$males.allages.avg.vote,
names = c("Female","Male"),
col = c("pink","blue"))
```

## How Does Movie Duration Affect a Movie's Success?

In the section below, we will look to see if there is any correlation between the duration
of movies and the average votes for the movie.

### Initial Model

```{r}
duration_lm <- lm(avg.vote~ duration)
summary(duration_lm)
plot(duration,avg.vote)
abline(duration_lm, col = 'red')
```

### Model after taking out some outlier values

```{r}
res_std <- rstandard(duration_lm)
USAMOVIES1$res_std = res_std
USAMOVIES2 <- USAMOVIES1[USAMOVIES1$res_std <= 2,]
USAMOVIES3 <- USAMOVIES2[USAMOVIES1$res_std >= -2,]
duration_lm2 <- lm(USAMOVIES3$avg.vote~ USAMOVIES3$duration)
summary(duration_lm2)
plot(USAMOVIES3$duration,USAMOVIES3$avg.vote)
abline(duration_lm2, col= 'red')
```

We can see from the summary above that the model and the variable for duration is 
significant when predicting the average vote.




